---
title: 'Lab 4: Tidying and Joining'
author: "Hannah Lo(halo-ds) and Yuma Anderson(yumaanderson)"

date: "10/24/2019"
output: html_document
---

```{r, include=FALSE}
library(dplyr)
library(plyr)
library(ggplot2)
library(tidyr)
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

## Part One: Cleaning Data

Changed the player names and opponent names to factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)

bio$Name <- as.factor(bio$Name)
```

Changed the offensive and defensive statistics to numerical, along with the 'Weight' column of Biography. 
```{r}
bio$Weight <- as.numeric(bio$Weight)

offClean <- offense %>% mutate_if(is.character, as.numeric)

defClean <- defense %>% mutate_if(is.character, as.numeric)
```

Separated the `Height` column into two columns, and made sure the newly created columns are numeric. After, we created one column that represents `Height`. Lastly, we dropped the columns 'Feet' and 'Inches' from biography.
```{r}
bio <- separate(data = bio, col = Height, into = c("Feet", "Inches"), sep = "-")
bio$Feet <- as.numeric(bio$Feet)
bio$Inches <- as.numeric(bio$Inches)
bio$HeightinInches <- (bio$Feet * 12 ) + bio$Inches
bioClean <- subset(bio, select = -c(Feet, Inches))
```

We have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. These are their following structures of these data frames:
```{r}
str(bioClean)

str(offClean)

str(defClean)
```


## Part Two: Tidying (A)

We reformated the `defClean` data frame into a tidy format using `gather` on the type of statistic (Tackles_Solo, Tackles_ASST, etc) resulting in a new column 'Statistic.'
```{r}
defClean = defClean %>% gather(key=Statistics, value=score,3:11)
defClean$Statistics <- as.factor(defClean$Statistics)
# making categories of defense statistics clear
defClean$Statistics = revalue(defClean$Statistics, c("Turnover_FF" ="Turnover_Forced_Fumble", "Turnover_FR"="Turnover_Fumble_Recovered","Turnover_INT"="Turnover_Interception"))
```

The following is a histogram for each defensive statistic. Turnovers by forced fumble, fumble recover, and interception are rare defensive skills. 
```{r}
defClean %>% ggplot(aes(score))+geom_histogram(binwidth = 1) +facet_wrap(~Statistics)+labs(title= 'Defensive Statistic Histogram')
```

ISU had a better defense against Kansas State than West Virginia. This is shown by a scatterplot with the x-axis being the number of solo tackles against West Virginia and the y-axis being that against Kansas State. 
```{r}
defClean %>% subset(Statistics == "Tackles_Solo") %>% spread(key=Opponent_Opponent, value = score) %>% select(`West Virginia`, `Kansas State`) %>% ggplot(aes(x=`West Virginia`,y=`Kansas State`)) + geom_point() + labs(title='Solo Tackles against')
```
      
    
## Part two: Tidying (B)

We separated the Hometown column into two columns by the city and state. 
```{r}
bioClean = separate(data = bioClean, col = Hometown, into = c("City", "State"), sep = ",")
bioClean[1:2,]
```

The following table shows how many players there are from each state:
```{r}
table(bioClean$State)
```


## Part three: Joining data frames
By completing a left join on the offensive data and player biography data, the following table was made showing all offensive player's hometowns.
```{r}
off_bio = left_join(offClean, bioClean, by = "Name")
off_bio %>% group_by(Name) %>% select(City,State) %>% distinct()
```

There is no relationship between the weight of a player and receiving yards. This is shown by the following scatterplot of the receiving yards by a player's weight: 
```{r message=FALSE, warning=FALSE}
off_bio %>% ggplot(aes(Weight,Receiving_YDS)) + geom_point(position = 'jitter', alpha= .5) +labs(title="Receiving yards by weight",y="Receiving Yards")
```

A question we can ask ourselves is what players are exclusively offensive? The following list shows players that solely play offense. 
```{r}
offClean %>% anti_join(defClean,by="Name") %>% select("Name") %>% distinct()
```





