# halo_ds 
# yumaanderson

## Loading the Data
```{r}
library(dplyr)
library(tidyr)
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```
```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

## Part One: Cleaning Data

1. The player names and opponent names are factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)

bio$Name <- as.factor(bio$Name)
```

2. All offensive and defensive statistics are numerical. The `Weight` column in Biography is also numeric. (Hint: look at `?mutate_at` if you want to save writing some code). 
```{r}
bio$Weight <- as.numeric(bio$Weight)

offClean <- offense %>% mutate_if(is.character, as.numeric)

defClean <- defense %>% mutate_if(is.character, as.numeric)
```


3. Separate the `Height` column into two columns. Make sure the newly created columns are numeric. Then create one column that represents `Height` (Hint: Use a unit that works well). Drop any irrelavant columns from this process. 
```{r}
bio <- separate(data = bio, col = Height, into = c("Feet", "Inches"), sep = "-")
bio$Feet <- as.numeric(bio$Feet)
bio$Inches <- as.numeric(bio$Inches)
bio$HeightinInches <- (bio$Feet * 12 ) + bio$Inches
bioClean <- subset(bio, select = -c(Feet, Inches))
```


4. By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames. **Make sure your data types make sense for each variable.**
```{r}
str(bioClean)

str(offClean)

str(defClean)
```


## Part Two: Tidying (A)

1. Reformat the `defClean` data frame into a tidy format using `gather`. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a **new key column named `Statistic`.** 
```{r}

```


2. Create a histogram for each defensive statistic (make sure to explore different binwidths). Use `facet_wrap` to draw all histograms in a single plot. What defensive skills are rare?
```{r}

```

3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense. 

      (**Hint**: From step 1 we have gathered our `defClean` into 4 columns. \
       Now we only want to focus on solo tackles (hint...`filter`). 
      Look at your data structure. What form do we need our data to have in order to easily answer this question? (hint...`spread`)). 
      
    
    
    
    
    
      
      
      